Temporality of the Victorian Hospital Patient Transfer Network

### 2025-06-30 reports/paper.qmd
yep week

---
title: "Temporality of the Victorian Hospital Patient Transfer Network"
format:
  plos-pdf:
    number-sections: false
    journal:
      id: ploscompbio
    include-in-header:
      text: |
        \usepackage{setspace}
        \doublespacing
    keep-tex: true
    fig-format: png
    fig-dpi: 300
author:
  - name: Michael J Lydeamore
    affiliations:
      - ref: ebs
      - ref: vicdh
    email: michael.lydeamore@monash.edu
    corresponding: true
  - name: David Wu
    affiliations:
      - ref: ebs
  - name: Tjibbe Donker
    affiliations:
      - ref: freiburg
  - name: Ben Cooper
    affiliations:
      - ref: oxford
  - name: Marion Easton
    affiliations:
      - ref: vicdh
  - name: Nicholas Geard
    affiliations:
      - ref: uom
  - name: Claire Gorrie
    affiliations:
      - ref: doherty
      - ref: mdu
  - name: Daneeta Hennessy
    affiliations:
      - ref: vicdh
  - name: Benjamin Howden
    affiliations:
      - ref: mdu
  - name: Anton Y Peleg
    affiliations:
      - ref: id
  - name: Charlie Higgs
    affiliations:
      - ref: mdu
  - name: Andrew Wilson
    affiliations:
      - ref: vicdh
  - name: Andrew J Stewardson
    affiliations:
      - ref: id

@@

# Methods

## Setting

Victoria, Australia, has a population of 7.0 million people [@ABS_state_populations]. Its hospital system consists of $N_x = 338$ healthcare facilities that vary in size, case mix and services. 
Admissions data was sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official].
In particular, data from 1 January 2011 to 9 December 2020 (3631 days) was used in this study. All Victorian hospital admissions during this study period are included. The total number of admission records is 26876787.
This data was aggregated into time-discretised temporal networks (described below), and exported from the database; further analysis as presented in this paper is performed on external machines with no access to identifiable protected health information.

Some exploratory analysis of this data was done in [@lydeamore_carbapenemase-producing_2024] using a static network framework.

## Network models of patient transfers for temporal fidelity

>>>>

---
title: "Temporality of the Victorian Hospital Patient Transfer Network"
format:
  plos-pdf:
    number-sections: false
    journal:
      id: ploscompbio
    include-in-header:
      text: |
        \usepackage{setspace}
        %\doublespacing
    keep-tex: true
    fig-format: png
    fig-dpi: 300
author:
  - name: Michael J Lydeamore
    affiliations:
      - ref: ebs
      - ref: vicdh
    email: michael.lydeamore@monash.edu
    corresponding: true
  - name: David Wu
    affiliations:
      - ref: ebs
  - name: Tjibbe Donker
    affiliations:
      - ref: freiburg
  - name: Ben Cooper
    affiliations:
      - ref: oxford
  - name: Marion Easton
    affiliations:
      - ref: vicdh
  - name: Nicholas Geard
    affiliations:
      - ref: uom
  - name: Claire Gorrie
    affiliations:
      - ref: doherty
      - ref: mdu
  - name: Daneeta Hennessy
    affiliations:
      - ref: vicdh
  - name: Benjamin Howden
    affiliations:
      - ref: mdu
  - name: Anton Y Peleg
    affiliations:
      - ref: id
  - name: Charlie Higgs
    affiliations:
      - ref: mdu
  - name: Andrew Wilson
    affiliations:
      - ref: vicdh
  - name: Andrew J Stewardson
    affiliations:
      - ref: id

@@

# Materials and Methods

We use admissions data sourced from the Victorian Admitted Episdoes Dataset (VAED) [@VAED_official]. This contains line-listed admissions for patients in Victoria, Australia, covering 338 healthcare facilities that vary in size, case mix and services.
the dataset contains 26,796,407 individual admissions across 5,471893 unique patients, from 1 January 2011 through to 6 December 2019 (3631 days).
Because there is no unique healthcare identifier in Victoria, Australia, linkage is performed on the VAED dataset by the Centre for Victorian Data Linkage, in order to build information on succesive admissions of patients. As a result of this linkage and hospital administrative procedures, there can be overlapping admissions, where a patient admits at a facility before being discharged at a prior facility. These overlaps are removed by constructing new admissions that are of minimal non-overlapping duration.
Some exploratory analysis of this data was done in [@lydeamore_carbapenemase-producing_2024] using a static network framework.

## Data Characterisation

To motivate the following methods, we first characterise temporal trends in the data.

Firstly, we examine the duration of stays

```{python}
#| label: fig-stay-duration
#| fig-cap: Distribution of duration of stays

stay_dur_df = pl.read_csv(root / "data/export/stay_duration_hist.csv")

fig, ax = plt.subplots()
sns.scatterplot(stay_dur_df, x='duration', y='count')
```
